---
title: "Assignment 3"
author: "Paola N. Kalil Rivera"
date: "`r format(Sys.time(), '%Y-%b-%d, %I:%M %p')`"
output: html_document
---

### __R Tutorial__
```{r}
library(ggplot2)
```

```{r}
data("trees")
head(trees)
```
##### __Q1.__<br>
__Create a histogram showing the distribution of Girth.__<br>
__a.__
```{r, message = FALSE}
ggplot(trees, aes(Girth)) +
  geom_histogram() +
  labs(title = "Distribution of Girth")
```

__b.__<br>
```{r, message = FALSE}
color = c("lightblue")
ggplot(trees, aes(Girth)) +
  geom_histogram(fill = color) +
  labs(title = "Distribution of Girth") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12)
  )
```

##### __Q2.__<br>
__Create a histogram showing the distribution of Volume.__<br>
__a.__<br>
```{r, message = FALSE}
ggplot(trees, aes(Volume)) +
  geom_histogram() +
  labs(title = "Distribution of Volume")
```

__b.__<br>
```{r, message = FALSE}
color = c("lightgreen")
ggplot(trees, aes(Volume)) +
  geom_histogram(fill = color) +
  labs(title = "Distribution of Volume") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12)
  )
```

##### __Q3.__<br>
__Create a scatterplot showing the relationship between Girth and Volume.__<br>
__a.__<br>
```{r}
ggplot(trees, aes(Girth, Volume)) +
  geom_point() +
  labs(title = "Relationship between Girth and Volume")
```

__b.__<br>
```{r}
ggplot(trees, aes(Girth, Volume)) +
  geom_point(color = "darkblue") +
  labs(title = "Relationship between Girth and Volume") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12)
  )
```

__c.__<br>
```{r, message = FALSE}
ggplot(trees, aes(Girth, Volume)) +
  geom_point(color = "darkblue") +
  labs(title = "Relationship between Girth and Volume") +
  geom_smooth(method = "lm", se = F, color = "darkred") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12)
  )
```





##### __Q4.__<br>
__Use the Shapiro–Wilk test to assess whether Volume is normally distributed. Report the p-value and state whether the data are normally distributed.__<br>
```{r, warning = FALSE}
TV <- aov(Volume ~ Volume, data = trees)
summary(TV)
resid_anova <- residuals(TV)
shapiro.test(resid_anova)
```
__Comment:__ Since the p-value (p = 0.003579) is less than 0.05, the null hypothesis is rejected and it is determined that the data do not follow a normal distribution.

##### __Q5.__<br>
__Repeat the Shapiro–Wilk test for Girth. Report the p-value and state whether Girth is normally distributed.__<br>
```{r, warning = FALSE}
TG <- aov(Girth ~ Girth, data = trees)
summary(TG)
resid_anova <- residuals(TG)
shapiro.test(resid_anova)
```
__Comment:__ Since the p-value (p = 0.08893) is greater than 0.05, the null hypothesis is not rejected and it is determined that the data do follow a normal distribution.