---
title: "Assignment 2"
author: "Paola N. Kalil Rivera"
date: "`r format(Sys.time(), '%Y-%b-%d, %I:%M %p')`"
output: html_document
---

### __R Tutorial__
```{r, message=FALSE}
library(dplyr)
```
```{r}
library(gapminder)
```
```{r}
data(gapminder)
head(gapminder)
```
##### __Q1.__<br>
##### What class is the gapminder dataset?<br>
```{r}
class(gapminder)
```
##### __Q2.__<br> 
##### Filter a section of gapminder with a life expectancy smaller than 29.<br>
```{r}
filter(gapminder, lifeExp < 29)
```
##### __Q3.__<br>
##### Filter the dataset for the country Rwanda and for years greater than 1979.<br>
```{r}
filter(gapminder, country == "Rwanda", year > 1979)
```
##### __Q4.__<br>
##### Create a new dataset from gapminder excluding the column gdpPercap.<br>
__a.__ Method 1
```{r}
Newgapminder <- gapminder |> 
  select(country,
         continent,
         year,
         lifeExp,
         pop)
Newgapminder
```
__b.__ Method 2
```{r}
Ngp <- gapminder |> 
  select(!(gdpPercap))
Ngp
```
__c.__ Method 3
```{r}
Ngp2 <- gapminder |> 
  select(-gdpPercap)
Ngp2
```
__Observations:__ Even tho I would get the same result, I wanted to try different variations of the function. Methods 2 and 3 decreased the amount of text required in the function.<br>

##### __Q5.__<br>
##### Arrange the dataset in ascending order of lifeExp.<br>
```{r}
Newgapminder |> arrange(lifeExp)
```
##### __Q6.__<br>
##### Arrange the dataset in descending order of lifeExp.<br>
```{r}
Newgapminder |> arrange(desc(lifeExp))
```
##### __Q7.__<br>
##### Create a new dataset with only the columns year and lifeExp.<br>
```{r}
Newgap <- gapminder |> 
  select(year,
         lifeExp)
Newgap
```
##### __Q8.__<br>
##### Filter the dataset for the country Cambodia and show only the columns year and lifeExp.<br>
```{r}
filter(gapminder, country == "Cambodia") |> 
select(year, lifeExp)
```
__Observations:__ There is more than one way to retrieve only the year and life expectancy columns from the filtered dataset.
At first, I forgot to put "Cambodia" in quotation marks, which caused an error. Also, by not including the |> after the filter, the command executed independently instead of being combined.<br>
```{r}
filter(gapminder, country == "Cambodia") |> 
select(-country, -continent, -pop, -gdpPercap)
```
##### __Q9.__<br>
##### Select only the countries in Africa.<br>
```{r}
filter(gapminder, continent == "Africa") |> 
  select(country)
```
##### __Q10.__<br>
##### Create a new column called lifeExp_months that shows life expectancy in months.<br>
__a.__ New column alongside all of the original columns.
```{r}
Newminder <- gapminder |> mutate(lifeExp_months = lifeExp * 12)
Newminder |> 
  select(lifeExp_months, lifeExp, everything())
```
__b.__ New column alongside the original life expectancy column.
```{r}
Newminder |> 
  select(lifeExp_months, lifeExp)
```